Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-16 nips_2018.tex
different introduction

Imagine that you are asked to edit some text,
and told that you should change the text ``Nancy FreeHafer'' to ``Dr. Nancy''.
From this example,
you likely infer that ``Pushmeet Kohli'' should be changed to ``Dr. Pushmeet'',
drawing upon your prior knowledge of text,
like that words are separated by spaces and that
one commonly prepends titles like ``Dr.''
In this work,
we consider the problem of
building agents that solve few-shot learning tasks like these,
and also the problem of acquiring
the prior knowledge necessary to quickly
solve these tasks (Figure~\ref{firstPageFigure}).
We think of
solutions to these tasks as being well represented by programs,
and so our problem can be stated as follows:
how should an agent learn to write programs?
We take inspiration from two sources:
(1) Good software engineers compose libraries
of reusable subroutines
that are shared across
related programming tasks.
Returning to Figure~\ref{firstPageFigure},
a good string processing library should support appending strings and splitting on spaces --
exactly the prior knowledge needed to solve the task in Figure~\ref{firstPageFigure}.
(2) Skilled human programmers
can quickly recognize what kinds of programming idioms and library routines would be useful for solving
the task at hand, even if they cannot instantly work out the details.
\begin{figure}[h]
  \begin{minipage}[c]{0.68\textwidth}
\hspace{-0.5cm}      \begin{tabular}{r|c|}\cmidrule{2-2}
  \sc{Task}&
  \begin{tabular}{rcl}
  Nancy FreeHafer&$\longrightarrow$& Dr. Nancy\\
% Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
 Pushmeet Kohli&$\longrightarrow$&???
  \end{tabular}\\\cmidrule{2-2}
  \sc{Program}&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}\\\cmidrule{2-2}
  \begin{tabular}{c}
    \sc{Library}\\
    \sc{(DSL)}
    \end{tabular}&\begin{tabular}{l}
    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y)}\\
    \phantom{$f_0($\code{a}$,$\code{b}$) \,=\, $\code{(}}\code{(cons x y)))}\\
    \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a)}\\
    \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}\\\cmidrule{2-2}
  \end{tabular}
  \end{minipage}%
  \begin{minipage}[c]{0.31\textwidth}
      \caption{\textbf{Task}: Few-shot learning problem. Model solves tasks by writing \textbf{program}s, and jointly learns a \textbf{library} of reusable subroutines that are shared across multiple tasks, called a \textbf{Domain Specific Language (DSL)}. Program writing is guided by a neural network trained jointly with the library.}\label{firstPageFigure}
    \end{minipage}
\end{figure}
We combine these two ideas into
a new algorithm called \systemEnding,
which takes as input a collection of
programming \textbf{tasks},
and then jointly
solves three problems:
(1) Writing programs that solve the tasks;
(2) Composing a library of
domain-specific
subroutines -- which allow the agent to
more compactly write programs in the domain,
and (3) Training a neural network
to recognize
which library components are
useful for which kinds of tasks.
Together, the library and neural net
encode the domain specific knowledge needed to
quickly write programs.

We call the learned library of subroutines a \textbf{Domain-Specific
  Language} (\textbf{DSL}), a term widely used in the program
synthesis community~\cite{polozov2015flashmeta}.
Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
The contribution of this work is
an algorithm for learning DSLs while jointly training a neural net to
search for programs in the DSL.

@@

Because any model may be encoded as a (deterministic or probabilistic) program,
we carefully delineate the scope of program learning problems considered here.
We think of \system as learning to solve the kinds of problems that humans
can solve relatively quickly -- once they acquire the relevant domain expertise.
These correspond to short programs -- if you have an expressive DSL.
Even with a good DSL, program search may be intractable, so
we amortize the cost of program search by training a neural network to
assist the search procedure.

Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding),
because it iterates between three different steps: an
\textbf{Explore} step uses the DSL to explore the space of programs,
searching for ones that solve the tasks; a \textbf{Compress} step
modifies the structure of the DSL by discovering regularities across programs found by
the previous Explore step; and a \textbf{Compile} step, which
improves the program search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}.
We call the neural net a \textbf{recognition model} (c.f. Hinton 1995~\cite{hinton1995wake}).
The learned DSL
distills commonalities across programs that solve tasks, helping
the agent solve related program induction problems. The neural
recognition model ensures that searching for programs remains tractable
even as the DSL (and hence the search space for programs) expands.

We apply \system to four domains:
list processing; FlashFill-style~\cite{gulwani2011automating} string editing;  symbolic regression; and turtle graphics~\cite{turtle}.
 For each of these we initially provide a generic
 set of programming primitives.
Our algorithm then discovers
its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{figure}
  \begin{tabular}{lcl}
    \toprule
    Domain&Example Task&Part of the learned DSL\\\midrule
    Lists&
    \begin{tabular}{rcl}
      \code{[7\, 0\, 2]}&$\longrightarrow$&\code{[7\, 0\, 2\, 4]} \\
      \code{[3\, 9]}&$\longrightarrow$&\code{[3\, 9\, 4]}
    \end{tabular}
    &
    \begin{tabular}{l}
      \code{(foldr nil (lambda (a b) (cons a b)))}\\
          \hspace{0.5cm} \emph{(appends lists)}
      \end{tabular}

@@

@@

To make this concrete we need to define what programs actually are and
what form $\probability[p |\mathcal{D},\theta]$ takes.
We represent programs as $\lambda$-calculus expressions.
$\lambda$-calculus is a formalism for expressing functional programs
that closely resembles the Lisp programming language.
$\lambda$-calculus includes variables, function application, and the ability to create new functions.
Throughout this paper we will write $\lambda$-calculus expressions in Lisp syntax.
Our programs are all strongly typed.
We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is
used in functional programming languages like OCaml and Haskell.
%% Type variables are always written using lowercase Greek letters
%% and we write $\alpha\to \beta$ to mean a function that takes an input of type $\alpha$
%% and returns something of type $\beta$.
%% We use the notation $p:\tau$ to mean that the $\lambda$-calculus expression $p$
%% has the type $\tau$.
%% For example, to describe the type of the identity function
%% we would say \code{(lambda (x) x)}$:\alpha\to \alpha$.
%% We say a type $\alpha$ \emph{unifies} with $\tau$ if every expression
%% $p:\alpha$ also satisfies $p:\tau$. Furthermore, the act of \emph{unifying}
%% a type $\alpha$ with $\tau$ is to introduce constraints on the type
%% variables of $\alpha$ to ensure that $\alpha$ unifies with $\tau$.
%% See Supplement for more detail on program representation.
We now define DSLs:

>>>>

Imagine that you are asked to write a program to edit some text, and
told that you should change the text ``Nancy FreeHafer'' to
``Dr. Nancy''.  From this example, you likely infer that ``Jane
Goodall'' should be ``Dr. Jane'', drawing upon your prior knowledge of
text, like that words are separated by spaces. For few-shot program learning problems
like these, the program synthesis and AI communities have had great
success engineering program learning systems,
not just for text editing~\cite{gulwani2011automating} but also for graphics
programs~\cite{ellis2017learning}, planning
procedures~\cite{devlin2017neural}, and many others.  However, the
success of these systems hinges upon having a carefully
hand-engineered \textbf{Domain Specific Language (DSL)}.  DSLs
impart prior knowledge of a domain
by providing a restricted set
of finely-tuned programming primitives: for
%that are finely tuned to the problems in the domain: for
text editing,
these are primitives like appending and splitting on characters.
In this work,
we consider the problem of
building agents that solve program learning tasks,
and also the problem of acquiring
the prior knowledge necessary to quickly
solve these tasks (Figure~\ref{firstPageFigure}).
Our solution is an algorithm that learns a DSL
while jointly training a neural network to write programs in that learned DSL.
\begin{figure}[h]
  \begin{minipage}[c]{0.68\textwidth}
\hspace{-0.5cm}      \begin{tabular}{r|c|}\cmidrule{2-2}
  \sc{Task}&
  \begin{tabular}{rcl}
  Nancy FreeHafer&$\longrightarrow$& Dr. Nancy\\
% Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
 Jane Goodall&$\longrightarrow$&???
  \end{tabular}\\\cmidrule{2-2}
  \sc{Program}&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}\\\cmidrule{2-2}
  \begin{tabular}{c}
    \sc{Library}\\
    \sc{(DSL)}
    \end{tabular}&\begin{tabular}{l}
    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y)}\\
    \phantom{$f_0($\code{a}$,$\code{b}$) \,=\, $\code{(}}\code{(cons x y)))}\\
    \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a)}\\
    \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}\\\cmidrule{2-2}
  \end{tabular}
  \end{minipage}%
  \begin{minipage}[c]{0.31\textwidth}
      \caption{\textbf{Task}: Few-shot learning problem. Model solves tasks by writing \textbf{program}s, and jointly learns a \textbf{library} of reusable subroutines that are shared across multiple tasks, called a \textbf{Domain Specific Language (DSL)}. Program writing is guided by a neural network trained jointly with the library.}\label{firstPageFigure}
    \end{minipage}
\end{figure}

%% We think of
%% solutions to these tasks as being well represented by programs,
%% and so our problem can be stated as follows:
%% how should an agent learn to write programs?
We take inspiration from two sources:
(1) Good software engineers compose libraries
of reusable subroutines
that are shared across
related programming tasks.
Returning to Figure~\ref{firstPageFigure},
a good string processing library should support appending strings and splitting on spaces --
exactly the prior knowledge needed to solve the task in Figure~\ref{firstPageFigure}.
(2) Skilled human programmers
can quickly recognize what kinds of programming idioms and library routines would be useful for solving
the task at hand, even if they cannot instantly work out the details.
We combine these two ideas into
an algorithm called \systemEnding,
which takes as input a collection of
programming \textbf{tasks},
and then jointly
solves three problems:
(1) Writing programs that solve the tasks;
(2) Composing a library of
domain-specific
subroutines -- which allow the agent to
more compactly write programs in the domain,
and (3) Training a neural network
to recognize
which library components are
useful for which kinds of tasks.
Together, the library and neural net
encode the domain specific knowledge needed to
quickly write programs.

@@

Our algorithm is called \textbf{Explore/Compress/Compile} (\systemEnding),
because it iterates between three different steps: an
\textbf{Explore} step uses the DSL and neural network to explore the space of programs,
searching for ones that solve the tasks; a \textbf{Compress} step
modifies the structure of the DSL by discovering regularities across programs found by
the previous Explore step; and a \textbf{Compile} step, which
improves the program search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}.
We call the neural net a \textbf{recognition model} (c.f. Hinton 1995~\cite{hinton1995wake}).
The learned DSL
distills commonalities across programs that solve tasks, helping
the agent solve related program induction problems. The neural
recognition model ensures that searching for programs remains tractable
even as the DSL (and hence the search space for programs) expands.

Because any model may be encoded as a (deterministic or probabilistic) program,
we carefully delineate the scope of problems considered here.
We think of \system as learning to solve the kinds of problems that humans
can solve relatively quickly -- once they acquire the relevant domain expertise.
These correspond to short programs -- if you have an expressive DSL.
Even with a good DSL, program search may be intractable, so
we amortize the cost of program search by training a neural network to
assist the search procedure.

We apply \system to four domains:
list processing; FlashFill-style~\cite{gulwani2011automating} string editing; and symbolic regression.
 For each of these we initially provide a generic
 set of programming primitives.
Our algorithm then discovers
its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{figure}
  \begin{tabular}{lcl}
    \toprule
    Domain&Example Task&Part of the learned DSL\\\midrule
    Lists&
    \begin{tabular}{rcl}
      \code{[7\, 0\, 2]}&$\longrightarrow$&\code{[7\, 0\, 2\, 4]} \\
      \code{[3\, 9]}&$\longrightarrow$&\code{[3\, 9\, 4]}
    \end{tabular}
    &
    \begin{tabular}{l}
      \code{(foldr nil (lambda (a b) (cons a b)))}\\
          \hspace{0.5cm} \emph{(appends lists)}
      \end{tabular}

@@

Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
which also learns a library of subroutines.
We were inspired by EC,
and go beyond it by giving a new algorithm for learning DSLs, as well as a way of combining DSL learning with
neurally guided program search.
In the experiments section,
we compared directly with EC,
showing empirical improvements.
The contribution of this work is
then
an effective algorithm for learning DSLs which also trains a neural net to
search for programs in the DSL.

@@

To make this concrete we need to define what programs actually are and
what form $\probability[p |\mathcal{D},\theta]$ takes.
We represent programs as $\lambda$-calculus expressions.
$\lambda$-calculus is a formalism for expressing functional programs
that closely resembles Lisp,
including variables, function application, and the ability to create new functions.
Throughout this paper we will write $\lambda$-calculus expressions in Lisp syntax.
Our programs are all strongly typed.
We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is
used in functional programming languages like OCaml and Haskell.
%% Type variables are always written using lowercase Greek letters
%% and we write $\alpha\to \beta$ to mean a function that takes an input of type $\alpha$
%% and returns something of type $\beta$.
%% We use the notation $p:\tau$ to mean that the $\lambda$-calculus expression $p$
%% has the type $\tau$.
%% For example, to describe the type of the identity function
%% we would say \code{(lambda (x) x)}$:\alpha\to \alpha$.
%% We say a type $\alpha$ \emph{unifies} with $\tau$ if every expression
%% $p:\alpha$ also satisfies $p:\tau$. Furthermore, the act of \emph{unifying}
%% a type $\alpha$ with $\tau$ is to introduce constraints on the type
%% variables of $\alpha$ to ensure that $\alpha$ unifies with $\tau$.
%% See Supplement for more detail on program representation.
We now define DSLs:
